---
title: "Untitled"
output: html_document
date: "2024-03-10"
---

```{r}
library(ISLR)
library(caret)
library(class)
library(gmodels)
library(dplyr)
library(e1071)
```

```{r}
Universal <- read.csv('C:/Users/rotim/Documents/R/UniversalBank.csv')
summary(Universal)
```

To create dummy variables from the Education
```{r}
Education<-as.factor(Universal$Education)
Personal.Loan<-as.factor(Universal$Personal.Loan)
education_df <- data.frame(Education_1 = as.character(levels(Education)))
dummy_variables <- model.matrix(~ Education - 1, data = education_df)
```


To combine the main dataset and the new dummy variables created
```{r}
Universal <- cbind(Universal, dummy_variables)
summary(Universal)
```

Use the createDataPartition function to partition the Universal dataset using the Income column.
```{r}
set.seed(123) 
Temp_data=createDataPartition(Universal$Personal.Loan, p=0.60, list=FALSE)
```
Training data
```{r}
Training=Universal[Temp_data,]
```
Validation data
```{r}
Validation=Universal[-Temp_data,]
View(Validation)
```


A. Create a pivot table for the training data with Online as a column variable,
```{r}
pivot_table <-table(Universal$CreditCard, Universal$Online, Universal$Personal.Loan)
```

Naming the pivot table
```{r}
rownames(pivot_table) <- c("CreditCard_0", "CreditCard_1")
colnames(pivot_table) <- c("Online_0", "Online_1") 
```

```{r}
pivot_table<- addmargins(pivot_table)
pivot_table
```

B Consider the task of classifying a customer who owns a bank credit card and is actively using online banking services.
We can view this easily from the second pivot table where Personal.Loan = 1.
Looking at that table, the intersection of Credit_Card_1 and Online_1,we have 82
```{r}
 prob <- 82/882
 prob
```

C. Create two separate pivot tables for the training data. 
Pivot table with Loan as rows and Online as columns
```{r}
pivot_table_1 <-table(Universal$Personal.Loan, Universal$Online)
```

Naming the pivot table
```{r}
rownames(pivot_table_1) <- c("Personal.Loan,_0", "Personal.Loan,_1")
colnames(pivot_table_1) <- c("Online_0", "Online_1")
```

```{r}
pivot_table_1<- addmargins(pivot_table_1)
pivot_table_1
```

Pivot table with Loan as rows and CreditCard as columns
```{r}
pivot_table_2 <-table(Universal$Personal.Loan, Universal$CreditCard)
```

Naming the pivot table
```{r}
rownames(pivot_table_2) <- c("Personal.Loan,_0", "Personal.Loan,_1")
colnames(pivot_table_2) <- c("CreditCard_0", "CreditCard_1")
```

```{r}
pivot_table_2<- addmargins(pivot_table_2)
pivot_table_2
```

Compute the following quantities 
1 P(CC = 1 | Loan = 1)
```{r}
p1 <- 143/480
p1
```

2 P(Online = 1 | Loan = 1)
```{r}
p2 <- 291/480
p2
```


3 P(Loan = 1)
```{r}
p3 <- 480/5000
p3
```

 
4 P(CC = 1 | Loan = 0)
```{r}
p4 <- 1327/4520
p4
```


 
5 P(Online = 1 | Loan = 0)
```{r}
p5 <- 2693/4520
p5
```

6 P(Loan = 0)
```{r}
p6 <- 4520/5000
p6
```

E. Use the quantities computed above to compute the naive Bayes probability P(Loan = 1 | CC = 1, Online = 1).

P(Loan = 1 | CC = 1, Online = 1).= P(CC=1∣Loan=1)×P(Online=1∣Loan=1)×P(Loan=1)/(P( CC=1) ,P(Online=1))

P(Online = 1)
```{r}
p7 <- 2984/5000
```
P(Creditcard = 1)
```{r}
p8 <- 1470/5000
```

```{r}
prob_1 <- (p1*p2*p3)/ (p7*p8)
prob_1
```

F
The naive bayes value is a more accurate estimate
G. Which of the entries in this table are needed for computing P(Loan = 1 | CC = 1, Online = 1)?
1 P(CC = 1 | Loan = 1)
2 P(Online = 1 | Loan = 1)
3 P(Loan = 1)
4 P(Online = 1)
5 P(Creditcard = 1)

```{r}
str(Universal)
```

Run naive Bayes on the data. Examine the model output on training data...
Naive Bayes classifier
```{r}
nb_model <-naiveBayes(Personal.Loan~  CreditCard+Online, data = Training)
```

```{r}
Target_client <- data.frame(Personal.Loan = 1,Online = 1, CreditCard = 1)
```

Predict the default status of test dataset
```{r}
Predicted_Test_labels <-predict(nb_model, Target_client, type="raw")
Predicted_Test_labels
```

The result from running the model gives the probability of a client who took a personal loan, has a credit card and is using online services is 0.11, while the result of the computation is 0.09881.



