---
title: "Untitled"
output: html_document
date: "2024-03-10"
---

```{r}
library(ISLR)
library(caret)
library(class)
library(gmodels)
library(dplyr)
library(e1071)
```

```{r}
Universal <- read.csv('C:/Users/rotim/Documents/R/UniversalBank.csv')
summary(Universal)
```

To create dummy variables from the Education
```{r}
Education<-as.factor(Universal$Education)
Personal.Loan<-as.factor(Universal$Personal.Loan)
education_df <- data.frame(Education_1 = as.character(levels(Education)))
dummy_variables <- model.matrix(~ Education - 1, data = education_df)
```


To combine the main dataset and the new dummy variables created
```{r}
Universal <- cbind(Universal, dummy_variables)
summary(Universal)
```

Use the createDataPartition function to partition the Universal dataset using the Income column.
```{r}
set.seed(123) 
Temp_data=createDataPartition(Universal$Personal.Loan, p=0.60, list=FALSE)
```
Training data
```{r}
Training=Universal[Temp_data,]
```
Validation data
```{r}
Validation=Universal[-Temp_data,]
View(Validation)
```


A. Create a pivot table for the training data with Online as a column variable,
```{r}
pivot_table <-table(Universal$CreditCard, Universal$Online, Universal$Personal.Loan)
```

Naming the pivot table
```{r}
rownames(pivot_table) <- c("CreditCard_0", "CreditCard_1")
colnames(pivot_table) <- c("Online_0", "Online_1") 
```

```{r}
pivot_table<- addmargins(pivot_table)
pivot_table
```

B Consider the task of classifying a customer who owns a bank credit card and is actively using online banking services.
We can view this easily from the second pivot table where Personal.Loan = 1.
Looking at that table, the intersection of Credit_Card_1 and Online_1,we have 82
```{r}
 prob <- 82/882
 prob
```

C. Create two separate pivot tables for the training data. 
Pivot table with Loan as rows and Online as columns
```{r}
pivot_table_1 <-table(Training$Personal.Loan, Training$Online)
```

Naming the pivot table
```{r}
rownames(pivot_table_1) <- c("Personal.Loan,_0", "Personal.Loan,_1")
colnames(pivot_table_1) <- c("Online_0", "Online_1")
```

```{r}
pivot_table_1<- addmargins(pivot_table_1)
pivot_table_1
```

Pivot table with Loan as rows and CreditCard as columns
```{r}
pivot_table_2 <-table(Training$Personal.Loan, Training$CreditCard)
```

Naming the pivot table
```{r}
rownames(pivot_table_2) <- c("Personal.Loan,_0", "Personal.Loan,_1")
colnames(pivot_table_2) <- c("CreditCard_0", "CreditCard_1")
```

```{r}
pivot_table_2<- addmargins(pivot_table_2)
pivot_table_2
```

Compute the following quantities 
1 P(CC = 1 | Loan = 1)
```{r}
p1 <- 91/480
p1
```

2 P(Online = 1 | Loan = 1)
```{r}
p2 <- 179/278
p2
```


3 P(Loan = 1)
```{r}
p3 <- 278/3000
p3
```

 
4 P(CC = 1 | Loan = 0)
```{r}
p4 <- 792/2722
p4
```


 
5 P(Online = 1 | Loan = 0)
```{r}
p5 <- 1620/2722
p5
```

6 P(Loan = 0)
```{r}
p6 <- 2722/3000
p6
```

E. Use the quantities computed above to compute the naive Bayes probability P(Loan = 1 | CC = 1, Online = 1).

P(Loan = 1 | CC = 1, Online = 1).= P(CC=1∣Loan=1)×P(Online=1∣Loan=1)×P(Loan=1)/(P( CC=1) ,P(Online=1))

P(Online = 1)
```{r}
p7 <- 1799/3000
```
P(Creditcard = 1)
```{r}
p8 <- 883/3000
```

```{r}
prob_1 <- (p1*p2*p3)/ (p7*p8)
prob_1
```

F
The naive bayes value is a more accurate estimate as it takes into account the individual probabilities of each variable, as opposed to the pivot table that looks at the joint probability.

G. Which of the entries in this table are needed for computing P(Loan = 1 | CC = 1, Online = 1)?
1 P(CC = 1 | Loan = 1)
2 P(Online = 1 | Loan = 1)
3 P(Loan = 1)
4 P(Online = 1)
5 P(Creditcard = 1)

```{r}
str(Universal)
```

Run naive Bayes on the data. Examine the model output on training data...
Naive Bayes classifier
```{r}
nb_model <-naiveBayes(Personal.Loan~  CreditCard+Online, data = Training)
```

```{r}                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  
nb_model
```

The result from running the model gives the probability of a client who took a personal loan, has a credit card and is using online services is 0.644, while the result of the computation in E is 0.0640.
