---
output:
  pdf_document: default
  html_document: default
---


2. Import the dataset into R

TAA petition Dataset obtained from the below website [https://www.dol.gov/agencies/eta/tradeact/data/petitions-determinations###]


```{r}
library(readxl)
Petition <- read_excel('C:/Users/rotim/Documents/R/PetitionData_Revised.xlsx')
```

3. Print out descriptive statistics for a selection of quantitative and categorical variables.
Descriptive statistics of quantitative variables Inv.Days & Est.Workers.Obj 
```{r}

summary(Petition$Inv.Days)
summary(Petition$Est.Workers.Obj)

```
Descriptive statistics of categorical variables State & City
```{r}
summary(Petition$State)
table(Petition$State)
prop.table(table(Petition$State))

summary(Petition$City)
table(Petition$City)
prop.table(table(Petition$City))
```

4. Transform at least one variable. It doesn't matter what the transformation is.
Transformation of variable SIC to integer and Zip to character
```{r}

 Petition <- read_excel('C:/Users/rotim/Documents/R/PetitionData_Revised.xlsx')
Transform_SIC <- as.integer(Petition$SIC)
Transform_SIC

Transform_Zip <- as.character(Petition$Zip)
Transform_Zip
```

5. Plot at least one quantitative variable, and one scatterplot

Bar graph of the State variable
```{r}

library("ggplot2")
ggplot(Petition, aes(x= State))+
  geom_bar() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Scatter plot of the Inv.Days and Est.Workers.Obj variables
```{r}
library("ggplot2")
ggplot(Petition, aes(x= Inv.Days, y=Est.Workers.Obj ))+
  geom_point() +
  theme_minimal()
```
