```{r}
library(ISLR)
library(caret)
library(class)
library(gmodels)
library(dplyr)
library(e1071)
library(cluster)
library(factoextra)
library(kernlab)
library(ggcorrplot)
library(broom)
```


```{r}
pharma_1 <- read.csv('C:/Users/rotim/Documents/R/Pharmaceuticals.csv')
```
To select only numeric variables for clustering
```{r}
pharma <- pharma_1[sapply(pharma_1, is.numeric)]
pharma <- as.data.frame(pharma)
summary(pharma)
```
To scale  the data which is important when clustering
```{r}
pharma <- scale(pharma)
```
First we observe the correlation between our features.
```{r}
pharma_corr<- cor(pharma)
```
```{r}
ggcorrplot(pharma_corr,
           outline.color = "grey50",
           lab = TRUE,
           hc.order = TRUE,
           type = "full")
```
From the correlation matrix, we see most of the features are moderately correlated, with the exception of the following variables which exhibited a high correlation of 0.7 and above between each other; (ROA-Market_cap,RoE-ROA, Net_profit_margin-ROA). Due to the significance of the variables in this analysis, we do not omit or replace them.

We make use of kmeans clustering as our dataset is small and we want to conduct a simple and yet effective clustering analysis.
To determine the optimal no of clusters using the fviz_nbclust() 
```{r}
fviz_nbclust(pharma, kmeans, method = "wss")
```
From the graph below we see the elbow point at 2, hence we use 2 clusters.
```{r}
set.seed(123)
cluster <- kmeans(pharma, centers = 2, nstart = 25)
```
```{r}
cluster_table <- table(cluster$cluster)
cluster_table
```
```{r}
fviz_cluster(cluster, data = pharma)
```

b
To see the centers of each cluster, from this we can explain different characteristics of the dataset
```{r}
cluster$centers
```
From the first cluster, we see the pharmaceuticals are mostly have a high Return on Equity (ROE)
 and Return on Assets (ROA), indicating they make decent Net income/earnings, visible in the fact
that they also have high Net profit margins. We also see these companies are not heavily leveraged
hence reducing their interest expense burden and allowing for a higher net profit margin.
In the second cluster we see the companies have a low market cap ROE and ROA. It is no surprise that 
these companies are heavily leveraged, hence will have a bigger interest burden.It is important to note
that these companies also have high revenue growth and have a positive beta, meaning they move
positively with the markets and do very well in a favourable economy.


Plot a dendrogram to view the hierarchical clusters
Exclude non-numeric variables
```{r}
pharma_heir <- pharma_1[, -c(1, 12, 13, 14)]
```
Convert the data to a data frame
```{r}
pharma_heir <- as.data.frame(pharma_heir)
```
Compute the distance between variables
```{r}
pharma_dt <- dist(pharma_heir, method = "euclidean")
```
Create the hierarchical clusters

```{r}
hclust_1 <- hclust(pharma_dt, method = "ward.D2") 
```
To enable the hierarchy cluster display the pharmacy names.
```{r}
pharmacy_names <- pharma_1[, 2]
plot(hclust_1, cex = 0.5, main = "Hierarchical Clustering Dendrogram", xlab = "Pharmacy", ylab = "Distance", labels = pharmacy_names)
```

```{r}
cluster_vector <- cluster$cluster
summary(pharma_1)
table_Exchange<- table(cluster_vector, pharma_1$Exchange)
table_Location<- table(cluster_vector, pharma_1$Location)
table_Median_Recommendation<- table(cluster_vector, pharma_1$Median_Recommendation)
```

```{r}
table_Exchange
table_Location
table_Median_Recommendation
```
From the table_Exchange we note that all pharmacies from the first clusters, which have higher than average ROE, ROA and high Net_profit_margins we noted earlier are listed on the NYSE, based in the US.Also, about 8 pharmacies from the second clusters are located on the NYSE, while the remaining 2 are listed on the AMEX and NASDAQ exchange.
From the table_Location we note that most pharmacies from the first clusters, are concentrated in the US with two in the Uk, while about 4 pharmacies from the second cluster were listed in the EURO region.
From the table_Median_Recommendation we note that about 9 pharmacies from the first cluster have a hold or buy recommendation as they are returns are positive. About 7  clusters from the
second cluster also have a hold or buy recommendation, this is understandable as they have a higher than average revenue growth rate and a positive beta.
Overall we see that pharmacies from the first cluster are concentrated in the US and are all listed on the NYSE, although most pharmacies from both clusters have a positive hold/buy Median_recommendation.


d
For the first cluster we name them as High Market_cap,ROE,ROA and Low-Leverage Firms due to the characteristics
mentioned in question b.
For the second cluster we name them as High Leverage,revenue growth and Low ROE,ROA firms.

```{r}
cluster_names <- c("High Market_cap, ROE, ROA and Low-Leverage Firms", 
                   "High Leverage, Revenue Growth and Low ROE, ROA Firms")
```
Assign cluster names to the cluster object
```{r}
names(cluster$cluster) <- cluster_names
names(cluster$cluster)
```
